### 2019-10-18 thesis/txt/0-introduction.tex
writing loads of stuff on models, semi-structured data in rw, example pictures of build logs

CI builds produce long and verbose build logs~\cite{beller2017oops}, stating the progress and results of the various steps within the build. The structure of these logs changes greatly from project to project~\cite{staahl2014modeling} as it highly depends on the tools and environment used.
Build logs are a valuable data source. First of all, for developers, who read them to analyze why their build failed or how different stages of the build process perform.
Second, they are valuable for researchers, who can harvest the information contained in the logs --- and their metadata --- to study the software engineering process of a project.
However, these two groups can only utilize the information within the build logs if they can adequately retrieve the information relevant to them.

There are different retrieval techniques used to retrieve information from CI build logs. Beller et al. use regular expressions to analyze the failure reasons of Ruby and Java Maven build logs from TravisCI~\cite{beller2017oops}.
In general, such regular expressions are developed by looking at a few exemplary build logs and updating them whenever new cases are introduced is a tedious and error-prone task~\cite{michael2019regexes}.
Vassallo et al. wrote a custom parser for Java Maven build logs to gather information for build repair hints~\cite{vassallo2018un-break}.
Recently, Amar et al. greatly reduced the number of build log lines for a developer to inspect by creating a diff between the logs from a failed and successful build~\cite{amar2019mining}.

Currently there is only anecdotal evidence on the performance of these techniques and when a technique should be preferred. Developers and researches have little support while choosing which technique to use for an retrieval task. The goal for this thesis is to investigate different information retrieval techniques and describe under which circumstances we recommend each of the techniques.

We aim to define a model to characterize the different retrieval techniques and answer what influences the suitability of a technique.
We support our assumption by implementing and evaluating three techniques:
\begin{itemize}
  \item regular expression synthesis by example using the Microsoft PROSE library (referred to as PBE),
  \item a common text similarity approach (referred to as TS) and
  \item simple keyword search (referred to as SKWS)
\end{itemize}
To evaluate these techniques we create the \emph{Failing Build Logs Data Set}, which encompasses about 800 log files from 80 repositories. Each log is labeled with the log part describing the reason a build failed, keywords to search for this log part and a categorization of the labeled log part according to its structural position within the log.

We aim to answer the following research questions:
\begin{itemize}
  \item[\textbf{RQ1:}] Which criteria influence the suitability of an information retrieval technique for CI build logs?
  \item[\textbf{RQ2:}] When are PBE, TS and SKWS suited to retrieve information from CI build logs?
  \item[\textbf{RQ2.1:}] How many examples do PBE, TS and SKWS need to perform best?
  \item[\textbf{RQ2.2:}] How accurate are the retrievals of PBE, TS and SKWS?
  \item[\textbf{RQ2.3:}] How structurally similar do the examples for PBE and TS need to be for the techniques to be applicable?
\end{itemize}

@@

\paragraph{Our work contributes:}
\begin{itemize}
  \item A model of the retrievable information in CI build logs
  \item A model to characterize information retrieval techniques from CI build logs
  \item A model to characterize use case scenarios for information retrieval from CI build logs
  \item A tool unifying several information retrieval techniques namely:
        \begin{itemize}
          \item our implementation of regular expression program synthesis using the Microsoft PROSE library (PBE),
          \item a common information retrieval approach using text similarity (TS), and
          \item a simple keyword search approach (SKWS)
        \end{itemize}
  \item The \emph{LogCollector}, a tool to gather build logs from Travis CI
  \item A validated data set of about 800 logs from failed Travis CI builds labeled with:
        \begin{itemize}
          \item the substring of the log describing the reason the build failed,
          \item keywords developers would use to search for these descriptions of the build failure reason, and
          \item a categorization of the retrievals according to their structural position within the build log
        \end{itemize}
  \item An evaluation of our model for the three implemented information retrieval techniques \todo{describe \emph{what} we look for with this evaluation}
  \item A extendable scheme supporting decision on (our three) information retrieval techniques relative to a given use case scenario
\end{itemize}

>>>>

CI builds produce long and verbose build logs~\cite{beller2017oops}, stating the progress and results of the various steps within the build. The structure of these logs changes greatly from project to project~\cite{staahl2014modeling} as it highly depends on the tools and environment used.
Build logs are a valuable data source. First of all, for developers, who read them to analyze why their build failed or how different stages of the build process perform\mention{ citation needed }.
Second, they are valuable for researchers, who can harvest the information contained in the logs --- and their metadata --- to study the software engineering process of a project\mention{ citation needed vassallo beller}.
However, these two groups can only utilize the information within the build logs if they can adequately retrieve the information relevant to them.

There are different techniques used to retrieve information from CI build logs. Beller et al. use regular expressions to analyze the failure reasons of Ruby and Java Maven build logs from TravisCI~\cite{beller2017oops}.
In general, such regular expressions are developed by looking at a few exemplary build logs and updating them whenever new cases are introduced is a tedious and error-prone task~\cite{michael2019regexes}.
Vassallo et al. wrote a custom parser for Java Maven build logs to gather information for build repair hints~\cite{vassallo2018un-break}.
Recently, Amar et al. greatly reduced the number of build log lines for a developer to inspect by creating a diff between the logs from a failed and successful build~\cite{amar2019mining}.

Currently there is only anecdotal evidence on the performance of these techniques and when a technique should be preferred.
Developers and researches have little support while choosing which technique to use for an retrieval task.
The goal for this thesis is to investigate different information retrieval techniques (IR techniques) and describe under which circumstances we recommend each of the techniques.

We aim to define a model to characterize the different IR techniques as well as the information retrievable from CI build logs.
For our first research question we analyze which criteria influence the suitability of an IR technique.
We support our assumption by implementing and evaluating three techniques:
\begin{itemize}
  \item regular expression synthesis by example using the Microsoft PROSE library (referred to as PBE),
  \item a common text similarity approach (referred to as TS) and
  \item simple keyword search (referred to as SKWS)
\end{itemize}
Our second research question asks in which cases PBE, TS and SKWS are suited to retrieve information from continuous integration build logs. RQ2 is refined sub-questions along the criteria resulting from RQ1 and compares their instantiations for the three techniques: how many examples the techniques need to perform best, the accuracy of the retrieved output and how diverse the configuring examples can be structurally.
\begin{simplebox}{Research Questions}
\begin{itemize}
  \item[\textbf{RQ1:}] Which criteria influence the suitability of an information retrieval technique for CI build logs?
  \item[\textbf{RQ2:}] When are PBE, TS and SKWS suited to retrieve information from CI build logs?
  \item[\textbf{RQ2.1:}] How many examples do PBE, TS and SKWS need to perform best?
  \item[\textbf{RQ2.2:}] How accurate are the retrievals of PBE, TS and SKWS?
  \item[\textbf{RQ2.3:}] How structurally similar do the examples for PBE and TS need to be for the techniques to be applicable?
\end{itemize}
\end{simplebox}
To evaluate PBE, TS and SKWS we create the \emph{Failing Build Logs Data Set}, which encompasses about 800 log files from 80 repositories. Each log is labeled with the log part describing the reason a build failed, keywords to search for this log part and a categorization of the labeled log part according to its structural position within the log.

@@

\paragraph{Our work contributes:}
\begin{itemize}
  \item A model of the retrievable information in CI build logs
  \item A model to characterize information retrieval techniques from CI build logs
  \item A model to characterize use case scenarios for information retrieval from CI build logs
  \item A tool unifying several information retrieval techniques namely:
        \begin{itemize}
          \item regular expression program synthesis from examples using the Microsoft PROSE library (PBE),
          \item a common information retrieval approach using text similarity (TS), and
          \item a simple keyword search approach (SKWS)
        \end{itemize}
  \item The \emph{LogCollector}, a tool to gather build logs from Travis CI
  \item A validated data set of about 800 logs from failed Travis CI builds labeled with:
        \begin{itemize}
          \item the substring of the log describing the reason the build failed,
          \item keywords developers would use to search for these descriptions of the build failure reason, and
          \item a categorization of the retrievals according to their structural position within the build log
        \end{itemize}
  \item An evaluation of our model for the three implemented information retrieval techniques \todo{describe \emph{what} we look for with this evaluation}
  \item A extendable scheme supporting decision on (our three) information retrieval techniques relative to a given use case scenario
\end{itemize}

### 2019-10-18 thesis/txt/1-related-work.tex
fixes from Taicos introduction review
writing loads of stuff on models, semi-structured data in rw, example pictures of build logs

\section{Production Log Analysis}
\mention{logpai and drain, [7] K. Fisher and D. Walker. The pads project: an overview. In ICDT,
2011.
[8] K. Fisher, D. Walker, K. Q. Zhu, and P. White. From dirt to shovels:
fully automatic tool generation from ad hoc data. In POPL, 2008.[23] Q. Xi and D. Walker. A context-free markup language for semi- structured text. In PLDI, pages 221–232, 2010.
 The PADS project [7] has en- abled simplification of ad hoc data processing tasks for programmers
accounts addresses split chairs awk banks companies countries hadoop horses instruments ls-l mgx namephone nozzle numbertext papers pldi12 pldi13 pop13 quotes speechbench techfest ucla-faculty users
accounts addresses split chairs awk banks companies countries hadoop horses instruments ls-l mgx namephone nozzle numbertext papers pldi12 pldi13 pop13 quotes speechbench techfest ucla-faculty users
abt amazon apple barnes bestbuy bigtray bol buy cameraword cnet cooking-bw dealtime drugstore ebay mgzoutlet mediaworld nthbutsw powells googlepdct yahooshop shopping shopzilla target tigerdirect venere
abt amazon apple barnes bestbuy bigtray bol buy cameraword cnet cooking-bw dealtime drugstore ebay mgzoutlet mediaworld nthbutsw powells googlepdct yahooshop shopping shopzilla target tigerdirect venere
hg_ex12 hg_ex18 hg_ex2 hg_ex26 hg_ex29 hg_ex3 hg_ex39 _h8d62ck1* 03PFMJOU* 2003Fall 64040 anrep9899* bali ch15_e compliance* DataDiction* deliverable* e_Bubble_* flip_usd5 Funded - F ge-revenues HOSPITAL* pwpSurvey* SOA4-YEAR* young_table
hg_ex12 hg_ex18 hg_ex2 hg_ex26 hg_ex29 hg_ex3 hg_ex39 _h8d62ck1* 03PFMJOU* 2003Fall 64040 anrep9899* bali ch15_e compliance* DataDiction* deliverable* e_Bubble_* flip_usd5 Funded - F ge-revenues HOSPITAL* pwpSurvey* SOA4-YEAR* young_table
seconds
# examples
seconds
examples
seconds
examples
by contributing along several dimensions: development of domain specific languages for describing text structure or data format, learn- ing algorithms for automatically inferring such formats [8], and a markup language to allow users to add simple annotations to enable more effective learning of text structure [23] While PADS supports parsing of entire files, FlashExtract allows users to extract only parts of the file thereby avoiding unnecessary complications. PADS’s learner only supports a fixed line-by-line chunking strategy to split the records; in contrast, FlashExtract can learn chunking (aka, struc- ture boundaries) from examples, making it suitable for extracting data fields and records that have arbitrary length (and might cross multiple lines). Finally, PADS primarily targets ad hoc text files. Although one can view webpages and spreadsheet as text files, it is unclear if the PADS learning algorithm can be adapted to work effectively for webpages and spreadsheets.
}

@@

\section{Information retrieval and Retrieval Techniques}
\review{Getting the general topic or conceptual information of a text is a common task in information retrieval from semi-structured text sources. Usually this is done by preprocessing the documents, transforming them to a term-by-document matrix. On the matrix we apply a similarity comparison like for example vector space models to calculate the similarity of the different documents to each other~\cite{panichella2016parameterizing}. For our use case, this could be applied by slicing the build logs into lines or small sections. Then similarity measures are used to compare the overall topics in these subparts to the topic of previously labelled logparts.
Instead of receiving a paragraph which is similar to a given query, our works focus more on obtaining specific pieces of texts through regular expressions.}

@@

\end{document}

>>>>

\section{Production Log Analysis}
\mention{logpai and drain, [7] K. Fisher and D. Walker. The pads project: an overview. In ICDT,
2011.
[8] K. Fisher, D. Walker, K. Q. Zhu, and P. White. From dirt to shovels:
fully automatic tool generation from ad hoc data. In POPL, 2008.[23] Q. Xi and D. Walker. A context-free markup language for semi- structured text. In PLDI, pages 221–232, 2010.
 The PADS project [7] has en- abled simplification of ad hoc data processing tasks for programmers
by contributing along several dimensions: development of domain specific languages for describing text structure or data format, learn- ing algorithms for automatically inferring such formats [8], and a markup language to allow users to add simple annotations to enable more effective learning of text structure [23] While PADS supports parsing of entire files, FlashExtract allows users to extract only parts of the file thereby avoiding unnecessary complications. PADS’s learner only supports a fixed line-by-line chunking strategy to split the records; in contrast, FlashExtract can learn chunking (aka, struc- ture boundaries) from examples, making it suitable for extracting data fields and records that have arbitrary length (and might cross multiple lines). Finally, PADS primarily targets ad hoc text files. Although one can view webpages and spreadsheet as text files, it is unclear if the PADS learning algorithm can be adapted to work effectively for webpages and spreadsheets.
}

@@

\section{Information Extraction and Retrieval Techniques}
\todo{explain difference here, reference pads and stuff}
\review{Getting the general tot p cic or conceptual information of a text is a common task in information retrieval from semi-structured text sources. Usually this is done by preprocessing the documents, transforming them to a term-by-document matrix. On the matrix we apply a similarity comparison like for example vector space models to calculate the similarity of the different documents to each other~\cite{panichella2016parameterizing}. For our use case, this could be applied by slicing the build logs into lines or small sections. Then similarity measures are used to compare the overall topics in these subparts to the topic of previously labelled logparts.
Instead of receiving a paragraph which is similar to a given query, our works focus more on obtaining specific pieces of texts through regular expressions.}

@@

\section{Semi-Structured Data}
\label{sec:rw-semi-structured-data}
Serge Abiteboul ``characterizes essential aspects of semi-structured data''~\cite{abiteboul1997querying} in the context of integrating data from various data sources.
The content of CI build logs is combined from several tools involved in the build process and several of Abiteboul's aspects can be applied to build logs.

The structure of build logs
\begin{itemize}
  \item is \textbf{implicit}, as we do not have access to explicit structuring elements or an explicit structure description and computation is required to obtain the structure which is present.
  \item is \textbf{irregular}, as changes in the build process or execution environment might change it without notice.
  We observed this in the logs we collected for the \emph{Failing Build Log Data Set}, where for the same repository and build configuration some logs had double new line characters without any noticeable explanation.
  You can see an example of this in Figure \ref{fig:log-4} and Figure \ref{fig:log-5}.
  \item is \textbf{partial}, some parts are highly structured by e.g. structure indicating special characters and others are unstructured, e.g. natural language text in error messages.
  \item can described with an \textbf{a-posteriori data guide} than an a-priori schema.
  There is no fixed specification on how build tools have to structure their output, more we afterwards extract structuring patterns in the produced output.
  \item has a \textbf{rapidly evolving} schema.
  Modifications in the build configuration change the tools involved in the build and therefore the composition of the build log.
\end{itemize}

Abiteboul proposes an approach to overlay the semi-structured data with a structured layer. The additional layer can answer queries and give access to the semi-structured data relevant to the query. In contrast to that, this thesis takes a look at techniques to gater a specific, pre-specified information without the need to parse, understand or estimate the whole structure of a log.
\end{document}

### 2019-10-18 thesis/txt/2-models.tex
writing loads of stuff on models, semi-structured data in rw, example pictures of build logs

\section{Characteristics of a Build Log}
The idea of CI is, to integrate new software changes fast and often to catch errors early~\cite{humble2010continuous}. After making a change, the developer commits and pushes it to a shared source code repository. Companies often use a specific CI server, e.g. Travis CI, linked to their source code repository. A CI build might then be triggered by a push on specific branches or after a pull request was created. When such a CI build is triggered, it typically runs through these stages:
\begin{itemize}
	\item Pulling in the new, changed version of the source code.
	\item Running static analysis tools~\cite{zampetti2017open}.
	\item Building the software, i.e. compiling and packaging it~\cite{phillips2014understanding}.
	\item Running automated tests.
\end{itemize}
However these are only \emph{typical} stages and there is a high variability in the CI build processes of different software projects \mention{citation needed, zampetti2017open?}. Some smaller projects might just use CI to make sure their code compiles as a minimal check before reviewing a pull request, other projects might have various stages of extensive automated testing.

Independent from the actual composition of the CI build process, most software tools used within a build will write out log messages to the console \mention{citation needed, paper about usages of console.log?}. Often, the console output is the only means to for a build tool to communicate with its users. Tools print progress updates, error and warning messages on the console.

The structure of this output is chosen by every tool themselves. Many have implicit or explicit structuring rules, some adhere to predefined standards \mention{phpunit, ruby, xuint output?}. However, when anlayzing build logs we might not have access to the exact build configuration (which tools are used in which order) or might not have access to a useable definition of the output sturcture of a specific tool. Therefore we describe build logs as semi-structured.

Semi-structured for build logs means for us that the log text is:
\begin{itemize}
	\item \textbf{implicitly stuructured}, we do not have access to explicit structuring elements or an explicit structure description
	\item \textbf{irregular}, the structure changes without notice \todo{we saw this in our data collection, random newlines}
\end{itemize}

\plan{a build consists of various steps, executed by different tools, these are often orchestrated by at least one overarching tool. // picture: overarching tool travis and it calls different tools which tehn output sequentiall, travis sometimes outputs something in between -> that is how the build log is made up // we have this structure though it is not always apparent from the build log / edges between the systems can look very different, mostly no agreed model, we: look at differen techniques to retrieve certain pieces of information \emph{without complete parsing the whole log in to a structure} -> we don't necesarrily want to understand the whole log sturcture just retrieve one specifi information relevant to us}

\section{Retrievable Information in Build Logs}
\begin{figure}[h]
  \centering
\includegraphics[page=2, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
  \caption{Information retrievable from build logs}
  \label{fig:build-log-information-draft}
\end{figure}
Continuous integration build logs contain a huge amount of information about the various stages in the CI build they correspond to. In this section we present our model to more precisely describe our notion of a information retrievable from a CI build log.

A \emph{build log information} for us has:
\begin{itemize}
	\item a \emph{textual representation} within the build log, more specifically a substring of the log string
\end{itemize}
With this build information items can be hierachically ordered by their textual representations containing each other.

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{img/mt-graphics-BuildLogInformation.pdf}
	\caption{Build Log Information}
	\label{fig:build-log-information}
\end{figure}
During our inital exploration \& data collection for the \emph{Failing Build Log Data Set}, wanted to get an impression about the information one would want to retrieve from a build log. Within the Travis CI build logs we found various different informations possibly interesting to be retrieved
\begin{itemize}
	\item \textbf{Travis Build Step} A Travis CI build consists of several build steps defined within the Travis CI configuration language. Within the build log each of these steps are framed by \lstinline{travis_fold:start:<build step name>} and \lstinline{travis_fold:end:<build step name>}.
	      A build step contains
	      \begin{itemize}
		      \item \textbf{Build Step Name} The string Travis uses to identify the step within start \& end statements.
		      \item \textbf{Build Step Output} The output generated during the build step. The textual representation of the \texttt{Build Step Output} is the substring between the start \& end statements.
	      \end{itemize}
	\item \textbf{Travis Timing} Travis can measure the time of specified sections of the build process. It represents those by \lstinline{travis_time:start:<timing section id>} and \lstinline{travis_time:end:<timing section id>:start=<start time>,finish=<finish time>,duration=<duration>}
\end{itemize}

\todo{differentiate Travis CI specific things, explain high variability for instantiations}
\todo{give at least one concrete example instantiation, best from a log in the data set, not too long so maybe it goes in the abstract?, picture/graphic with marked parts here?}

\section{Characteristics of Information Retrieval Techniques}
\begin{figure}[h]
	\centering
	\includegraphics[page=3, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
	\caption{Model for an information retrieval technique}
	\label{fig:model-ie-technique}
\end{figure}
For this thesis we want to evaluate different information retrieval techniques in their application to CI build logs. We focus on techniques that do not attempt or require to parse the structure of a whole build log, but rather focus on extracting just one specific information
A build log information retrieval technique (BLIR technique) consumes a build log (text file) and produces a certain output (text). \todo{make picture, output can be contained in log or not, can be continuous substring or selected lines etc.}.
\begin{itemize}
	\item \textbf{Target}{Extractable Information} each retrieval technique \todo{instantiation!} targets a specific information
	\item \textbf{Setup Overhead}{Time}
	\item \textbf{Performance}{Performance} split into learning performance
	\item \textbf{Configuration Scope} for one repository/project, a language/logkind, global
\end{itemize}

@@

\section{Information Retrieval Tasks}
\begin{figure}[h]
	\centering
	\includegraphics[page=4, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
	\caption{Model for information extraction task}
	\label{fig:modelt-ie-task}
\end{figure}
aims to describe an information retrieval task of a developer or a researcher.
\begin{itemize}
	\item \todo{describe all classes}
	\item{Setup Overhead}{Time}
	\item{Performance}{Performance} split into learning performance
\end{itemize}
\todo{give 2-3 concrete example instantiations here}

@@

\todo{explanation of process, the way our tool does it, instantiate model of retrieval technique, probably no additional theoretical stuff really needed here?, describe regex programs}

\section{Text Similarity}
\begin{figure}[h]
  \centering
\includegraphics[page=3, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
  \caption{Extracting information using text similarity}
  \label{fig:text-similarity-explanation}
\end{figure}
\paragraph{Overview}
Information retrieval techniques are more and more prevalently used to extract textual information from unstructured textual sources~\mention{stuff annibale is referencing?}. We describe CI build logs as semi-structured, 
\todo{explanation of process, the way our tool does it, instantiate model of retrieval technique, probably no additional theoretical stuff really needed here?}

\section{Keyword Search}
\begin{figure}[h]
  \centering
\includegraphics[page=4, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
  \caption{Extracting information using keyword search}
  \label{fig:keword-search-explanation}
\end{figure}
\todo{explanation of process, the way our tool does it, instantiate model of retrieval technique, probably no additional theoretical stuff really needed here?}

>>>>

\section{Characteristics of a Build Log}
The idea of CI is, to integrate new software changes fast and often to catch errors early~\cite{humble2010continuous}.
After making a change, the developer commits and pushes it to a shared source code repository. Companies often use a specific CI server, e.g. Travis CI, linked to their source code repository.
A CI build might then be triggered by a push on specific branches or after a pull request was created.
When such a CI build is triggered, it typically runs through these stages:

\begin{itemize}
	\item \todo{look for more in modelling paper zampetti}
	\item Pulling the new, changed version of the source code onto the server.
	\item Running static analysis tools~\cite{zampetti2017open}.
	\item Building the software, i.e. compiling and packaging it~\cite{phillips2014understanding}.
	\item Running automated tests~\cite{beller2017oops}.
\end{itemize}

However these are only \emph{typical} stages and there is a high variability in the CI build processes of different software projects \mention{citation needed, zampetti2017open?}.
Some smaller projects might use CI to just make sure their code compiles as a minimal check before reviewing a pull request.
Other projects might have various stages of extensive automated testing.

Most software tools used within a build will write out log messages to the console \mention{citation needed, paper about usages of console.log?}.
To communicate progress updates, error and warning messages to the user.
This combined output is referred to as build log.
The structure of their output is chosen by every tool themselves.
Many have implicit or explicit structuring rules, some adhere to predefined standards \mention{phpunit, ruby, xuint output?}.
\todo{figure X} shows how different tools contribute to the whole build log.

When analyzing build logs we might not have access to the exact build configuration, describing which tools are used in which order, or might not have access to a useable definition of the output structure of a specific tool.
As the structure of a build log is not known to us, we describe build logs as semi-structured. Section \ref{sec:rw-semi-structured-data} describes different properties of semi-structured data.

\section{Retrievable Information in Build Logs}
\begin{figure}[h]
  \centering
\includegraphics[page=2, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
  \caption{Information retrievable from build logs}
  \label{fig:build-log-information-draft}
\end{figure}
Continuous integration build logs contain a huge amount of information about the various stages in the CI build they correspond to.
In this section we present a model, which more precisely describes our notion of a information retrievable from a CI build log.

The central class of our model is the \emph{build log information} (BLI), representing an piece of information possibly retrievable from a build log.
\emph{Possibly}, because a build log information is not necessarily present in every build log.
If it is present, we call it a \emph{build log information instantiation} (BLII), which always relates to a specific build log, the one it is appearing in.
\todo{add to model, build, build log, bulidloginformationinstantiation as association name to the type build log information}
Each BLII for us has a \emph{textual representation} within the build log.
This textual representation is always a substring of the log text.
With this build log information \todo{akward class names like BuildLogInformation? s would be possible then} can be hierarchically ordered by their textual representations containing each other.
As this is an ad-hoc and a-posteriori structuring schema, as we describe in Section \ref{sec:rw-semi-structured-data}, the model only presents hierarchical ordering or containment for build log informations whose structure we definetely know.
Most build log information types can appear in various hierarchical arrangements and therefore not contained in any other type in the model.

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{img/mt-graphics-BuildLogInformation.pdf}
	\caption{Build Log Information}
	\label{fig:build-log-information}
\end{figure}
During our initial exploration and log data collection for the \emph{Failing Build Log Data Set}, we collected a broad set of build logs from 29 languages and 87 repositories. We inspected them to get an impression about the information one would want to retrieve from a build log. Within the build logs from Travis CI we found various different information types possibly interesting to be retrieved:

\begin{itemize}
	\item \textbf{Travis Build Step} A Travis CI build consists of several build steps defined within the Travis CI configuration language. Within the build log each of these steps are framed by \lstinline{travis_fold:start:<build step name>} and \\ \lstinline{travis_fold:end:<build step name>}.
	These build steps could be categorized, for example into \textbf{Test Step}, \textbf{Linter Step}, \textbf{Travis Worker}, \textbf{Git Fetch}, or \textbf{Sudoers} are some that we encountered in the logs.
	A build step contains
	\begin{itemize}
		\item \textbf{Build Step Name} The string Travis CI uses to identify the step within start and end statements.
		\item \textbf{Build Step Output} The output generated during the build step. The textual representation of the \texttt{Build Step Output} Instantiation is the substring between the start and end statements.
	\end{itemize}
	You can see an example of a Travis Build Step and its components within a build log in Figure \ref{fig:log-1}.

	\item \textbf{Non-Travis Build Step??}
	\item 
	\item \textbf{Travis Timing} Travis can measure the time of specified sections of the build process.
	You can see an example of a Travis Build Step and its components within a build log in Figure \ref{fig:log-1}.
	%It represents those by \lstinline{travis_time:start:<timing section id>} and \\ \lstinline{travis_time:end:<timing section id>:start=<start time>,finish=<finish time>,duration=<duration>}

	\item \textbf{Travis Worker} Travis CI mentions for each build which machine is executing it.
	Retrieving this information from multiple build logs could for example help visualize the impact of the build server assignment algorithm.
	The Travis Worker is a good example to show that the same BLI can have different textual representations in different build logs.
	You can see examples of Travis Worker instantiations in Figure \ref{fig:log-0} and Figure \ref{fig:log-2}.

	\item \textbf{Travis System Info} At the beginning of each log Travis CI describes the tech stack of the server executing the build.
	The system info could be retrieved from both failing and successful logs and compared to identify if a failure is possibly based on the execution environment.
	You can see an example of a Travis System Info instantiation within a build log in Figure \ref{fig:log-0}.

	\item \textbf{Triggering Command} Travis CI logs the commands it uses to call certain tools.
	These come from the \texttt{travis.yml} configuring the build.
	Retrieving this information from the log could for example be useful when reverse engineering the build configuration.
	You can see an example of a Triggering Command instantiation within a build log in Figure \ref{fig:log-3}.

	\item \textbf{Exit Code} Travis CI prints all exit codes of commands.
	This information could be used to fill an overview of the build steps and why they failed.
	You can see an example of an Exit Code instantiation within a build log in Figure \ref{fig:log-3}.

	\item \textbf{Error Message} Various tools involved in the build process might output messages of errors that occurred during their execution.
	Retrieving these and showing them to developers could help them understand quicker why the build failed.
	You can see an example of and Error Message instantiation within a build log in Figure \ref{fig:log-3}.

	\item \textbf{Warning Message} In addition to errors, tools also print warning messages.
	These could be collected, counted and presented to developers, encouraging them to resolve them in order to improve their software.
	You can see examples of WarningMessage instantiations within build logs in Figure \ref{fig:log-4} and Figure \ref{fig:log-5}.

	\item \todo{name, explanation, ref to example image, example what to be used for}
\end{itemize}

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth, clip]{img/log0.png}
	\caption{Excerpt from a build log showing textual representations of a long Travis Worker instantiation and a Travis System Info instantiation}
	\label{fig:log-0}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth, clip]{img/log1.png}
	\caption{Excerpt from a build log showing textual representations of a Travis Timing instantiation and a Travis Build Step, containing the Build Step Name and the Build Step Output}
	\label{fig:log-1}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth, clip]{img/log2.png}
	\caption{Excerpt from a build log showing textual representations of a short Travis Worker instantiation}
	\label{fig:log-2}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth, clip]{img/log3.png}
	\caption{Excerpt from a build log showing textual representations of Error Message, Exit Code and Triggering Command instantiations}
	\label{fig:log-3}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth, clip]{img/log4.png}
	\caption{Excerpt from a build log showing textual representations of Warning Message instantiations}
	\label{fig:log-4}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth, clip]{img/log5.png}
	\caption{Excerpt from a build log showing textual representations of Warning Message instantiations}
	\label{fig:log-5}
\end{figure}

\section{Characteristics of Information Retrieval Techniques}
\begin{figure}[h]
	\centering
	\includegraphics[page=3, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
	\caption{Model for an information retrieval technique}
	\label{fig:model-ie-technique}
\end{figure}
For this thesis we want to evaluate different information retrieval techniques in their application to CI build logs. We focus on techniques that do not attempt or require to parse the structure of a whole build log, but rather focus on extracting just one specific information
A build log information retrieval technique (BLIR technique) consumes a build log (text file) and produces a certain output (text). \todo{make picture, output can be contained in log or not, can be continuous substring or selected lines etc.}.
\begin{itemize}
	\item \textbf{Target}{Extractable Information} each retrieval technique \todo{instantiation!} targets a specific information
	\item \textbf{Setup Overhead} {Time}
	\item \textbf{Performance} {Performance} split into learning performance
	\item \textbf{Configuration Scope} for one repository/project, a language/logkind, global
	\item \textbf{Granularity} of extraction, i.e. the smallest extractable text piece
\end{itemize}

@@

\section{Information Retrieval Tasks}
\begin{figure}[h]
	\centering
	\includegraphics[page=4, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
	\caption{Model for information extraction task}
	\label{fig:modelt-ie-task}
\end{figure}
aims to describe an information retrieval task of a developer or a researcher.
\begin{itemize}
	\item \todo{describe all classes}
	\item{Setup Overhead}{Time}
	\item{Performance}{Performance} split into learning performance
	\item 
\end{itemize}
\todo{give 2-3 concrete example instantiations here}

@@

\section{Text Similarity}
\begin{figure}[h]
  \centering
\includegraphics[page=3, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
  \caption{Extracting information using text similarity}
  \label{fig:text-similarity-explanation}
\end{figure}
\paragraph{Overview}
Text Similarity approaches are more and more used to filter unstructured textual software artifacts. \mention{cite papers that do exactily this. also lda possbible here}
One common and simple technique the Vector Space Model \mention{citation needed, for VSM and that it is common maybe?}.
We investigate when it is a suitable technique to retrieve information from build logs.

\paragraph{Configuration / Input / Output}
To configure a retrieval though text similarity we chose to use the same concept of I/O examples as for PBE.
The output strings of a given example set define our search query.
A retrieval run then takes as input a log file and outputs the parts most similar to the search query.

\paragraph{Similariy Calculation}
The TS retrieval technique works on a granularity of lines.
First, the search query is processed.
The output strings of the example set configuring the technique are split into single lines.
For each line we apply preprocessing: Splitting it into tokens, in our case words.
%Commonly text similarity techniques employ further normalization like removing special characters and stop words.
%We chose to keep the build log lines as complete as possible, as we often saw special characters and words identifying an interesting area.
Then we build a document-term-frequency matrix over the lines of the output strings and prune very often or very rarely appearing words.
Next, the algorithm applies term-frequency inverse document frequency, a best practice for natural language queries~\cite{lee1997document}.
To retrieve the desired information from a build log, we parse the whole text and process it in the same way.
The algorithm calculates the cosine simlarity~\mention{citation?} to compare each line of the build log with each line of the search query.
After summing up the similarities of each build log line to all the query lines, we sort the build log lines in decreasing similarity.
The average number of lines in the I/O example output strings determines how many of the most similar lines are returned as the output of the retrieval run.

\paragraph{Instantiate Retrieval Technique Model}

%Information retrieval techniques are more and more prevalently used to extract textual information from unstructured textual sources~\mention{stuff annibale is referencing?}. We describe CI build logs as semi-structured,

\section{Keyword Search}
\begin{figure}[h]
  \centering
\includegraphics[page=4, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
  \caption{Extracting information using keyword search}
  \label{fig:keword-search-explanation}
\end{figure}

\paragraph{Overview}
When we are looking for a specific information within a great amount of unstructured information, a first ad-hoc approach is to search for keywords in the text we assume to be near the information we are looking for.
This was one of the most common approaches we took when searching for the build failure reason within a log when creating our data set.
As this is a technique readily available in many tools developers use to view build logs, we want to study when such a simple search for keywords is recommendable for retrieving information from CI build logs.

\paragraph{Configuration / Input / Output}
SKWS is configured by giving a set of keywords.
For a retrieval run we then take a whole build log file as input and search for all exact occurrences of the keywords.
As keywords are often not directly describing the desired information but rather adjacent to the desired information, this technique not only retrieves the line with the keyword, but only the context around it.
We chose a context of 10 lines above and below the keyword appearance.

\paragraph{Instantiation of Model}

\todo{explanation of process, the way our tool does it, instantiate model of retrieval technique, probably no additional theoretical stuff really needed here?}
